---
title: "Time Series Cross-Correlation"
author: "Deepak Bastola"
date: "January 3, 2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
header-includes:
- \usepackage{bbm}
- \usepackage{amsmath}
---

+ VAR(1) model: 

\[
Y_t = \Phi \cdot Y_{t-1} + \epsilon_t, \qquad \text{where} \quad \epsilon_t \sim N(0, \Omega)
\]

+ Consider estimating $\bar{Y}_n$. The Yule- Walker Equations are,
$$
\begin{split}
\Gamma (0) &= \Phi \cdot \Gamma(1)' + \Omega \\
\Gamma (k) &= \Phi \cdot \Gamma(k-1), \qquad k = 1,2,3,......
\end{split}
$$

+ Wold Decomposition:
$$
vec(\Gamma(0)) = (I_{p^2} - \Phi \otimes \Phi)^{-1} \cdot vec(\Omega)
$$

+ Var-Correlation Decomposition:

$$
\begin{split}
D &= diag(\Gamma(0)) \\
R(h) &= D^{-\frac{1}{2}} \cdot \Gamma(h) \cdot D^{-\frac{1}{2}} 
\end{split}
$$

+ Let $V = \Gamma(0)$, then,
$$
\Sigma = (I_p - \Phi)^{-1} \cdot V + V (I_p - \Phi)^{-1} - V
$$

+ From Markov Chain CLT, we have,

\begin{align*}
\Sigma = \text{Var}_{\pi}(Y_1) + \sum_{k=1}^{\infty}[\text{Cov}_F(Y_1, Y_{1+k}) + \text{Cov}_F(Y_1,Y_{1+k})^T]. 
\end{align*}

+ MA($\infty$) representation of VAR(1) model,

$$
\begin{split}
Y_t &= \Phi \cdot Y_{t-1} + \epsilon_t\\
(I - \Phi L)Y_t &= \epsilon_t \\
Y_t &= (I - \Phi L)^{-1} \cdot \epsilon_t \\
Y_t &= \sum_{q = 0}^{\infty} \Phi_q \cdot \epsilon_{t-q}
\end{split}
$$

+ Then,

$$
\begin{split}
Cov(Y_t) = Cov\bigg(\sum_{q = 0}^{\infty} \Phi_q \cdot \epsilon_{t-q}\bigg) = \sum_{q = 0}^{\infty} \Phi_q \Omega \Phi_q^T 
\end{split}
$$

+ The correlation between $Y_i$ and $Y_j$ can be simply obtained by the transformation,

$$
R_{Y,ij} = \frac{\sigma_{Y,ij}}{\sqrt{\sigma_{Y,ii} \cdot \sigma_{Y,jj}}}
$$

+ To check the convergence rate of $R(h)$, we can do the following computation:

$$
\begin{split}
\Gamma(h) &= \Phi \Gamma_{h-1} = \Phi^h \Gamma_0 \\
R(h) &=  D^{-\frac{1}{2}} \cdot \Gamma(h) \cdot D^{-\frac{1}{2}} \\
&= D^{-\frac{1}{2}} \cdot \Phi\Gamma(h-1) \cdot D^{-\frac{1}{2}}\\
&= D^{-\frac{1}{2}} \cdot \Phi \cdot D^{-\frac{1}{2}}\cdot D^{\frac{1}{2}} \cdot \Gamma(h-1) \cdot D^{-\frac{1}{2}}\\
&= \Psi \cdot R(h-1) \\
&= \Psi^h \cdot R(0)
\end{split}
$$

where $\Psi = D^{-\frac{1}{2}} \cdot \Phi \cdot D^{-\frac{1}{2}}$. It can be seen that the correlation decay follows a power law.


# Some Examples


```{r, message=FALSE, warning=FALSE}
set.seed(123)
library(mAr)
library(Matrix)
library(mcmcse)
library(matrixcalc)
library(coda)
library(Matrix)

```

#

1. Bivariate VAR(1) with general correlation structure ($\Phi$).

```{r, message=FALSE, warning=FALSE}
p=2
n = 1000
omega <- diag(2)
Phi <- matrix(c(0.7, 0.1, 0.1, 0.5),2,2)
scratch <- diag(p^2) - kronecker(Phi,Phi)
V.s <- solve(scratch)%*%vec(omega)

Tau0 <- matrix(V.s, nrow = p, byrow = TRUE)
Tau0

D <- diag(diag(Tau0))
Sigma <- solve(diag(p)-Phi)%*%Tau0 + Tau0%*%solve(diag(p)-Phi) -Tau0
Sigma.d <- diag(diag(Sigma))
R.main <- solve(sqrt(Sigma.d))%*%Sigma%*%solve(sqrt(Sigma.d))


main <- solve(sqrt(D))%*%Sigma%*%solve(sqrt(D))

Tau1 <- Phi%*%Tau0
Tau2 <- Phi%*%Tau1
Tau3 <- Phi%*%Tau2
Tau4 <- Phi%*%Tau3
Tau5 <- Phi%*%Tau4


Rho0 <- solve(sqrt(D))%*%Tau0%*%solve(sqrt(D))

Rho1 <- solve(sqrt(D))%*%Tau1%*%solve(sqrt(D)) 
Rho2 <- solve(sqrt(D))%*%Tau2%*%solve(sqrt(D)) 
Rho3 <- solve(sqrt(D))%*%Tau3%*%solve(sqrt(D)) 
Rho4 <- solve(sqrt(D))%*%Tau4%*%solve(sqrt(D)) 
Rho5 <- solve(sqrt(D))%*%Tau5%*%solve(sqrt(D))

psi.h <- solve(sqrt(D))%*%Phi%*%(sqrt(D))

# Markov Chain
chain <- as.matrix(mAr.sim(rep(0,p), Phi, omega, N = n))

aa <- ccf(chain[,1], chain[,2], type = "correlation", plot = TRUE, lag.max = 100)
```

#
There is significant cross-correlation at large lags. There should be some tests in the literature to say if the cross-correlation decays geometrically, or follows some power law. 

#

```{r}
Tau1 <- Phi%*%Tau0
Tau2 <- Phi%*%Tau1
Tau3 <- Phi%*%Tau2
Tau4 <- Phi%*%Tau3
Tau5 <- Phi%*%Tau4

D <- diag(diag(Tau0));D

Rho0 <- solve(sqrt(D))%*%Tau0%*%solve(sqrt(D)) ;Rho0
Rho1 <- solve(sqrt(D))%*%Tau1%*%solve(sqrt(D)) ;Rho1
Rho2 <- solve(sqrt(D))%*%Tau2%*%solve(sqrt(D)) ;Rho2
Rho3 <- solve(sqrt(D))%*%Tau3%*%solve(sqrt(D)) ;Rho3
Rho4 <- solve(sqrt(D))%*%Tau4%*%solve(sqrt(D)) ;Rho4
Rho5 <- solve(sqrt(D))%*%Tau5%*%solve(sqrt(D)) ;Rho5

```

#

2. If we model the VAR(1) as VMA process then we can get few non-zero cross-correlations. I don't know if this helps.

#

```{r, message=FALSE, warning=FALSE}
# Example 2: 
p=2
omega <- matrix(c(1,0.1,0.1,1),2,2)
Phi <- matrix(c(0.5, 0, 0, 0.2),2,2)
scratch <- diag(p^2) - kronecker(Phi,Phi)
V.s <- solve(scratch)%*%vec(omega)

Tau0 <- matrix(V.s, nrow = p, byrow = TRUE)
Sigma <- solve(diag(p)-Phi)%*%Tau0 + Tau0%*%solve(diag(p)-Phi) -Tau0


# Markov Chain
chain <- as.matrix(mAr.sim(rep(0,p), Phi, omega, N = n))
aa <- ccf(chain[,1], chain[,2], type = "correlation", plot = TRUE, lag.max = 100)

```

#

There are only 2 significant cross-correlation terms.

#

3. I can extend this to higher dimensions.

```{r, message=FALSE, warning=FALSE}
p=3
omega <- matrix(c(1,0,0,0,1,0,0,0,1),3,3)
Phi <- matrix(c(0.3,0.3,0.3,0.6,0,0,0.8,0.6,0),3,3)

scratch <- diag(p^2) - kronecker(Phi,Phi)
V.s <- solve(scratch)%*%vec(omega)
Tau0 <- matrix(V.s, nrow = p, byrow = TRUE)

# Markov Chain
chain <- as.matrix(mAr.sim(rep(0,p), Phi, omega, N = n))

par(mfrow = c(1,2))
aa <- ccf(chain[,1], chain[,2], type = "correlation", plot = TRUE, lag.max = 100)
bb <- ccf(chain[,1], chain[,3], type = "correlation", plot = TRUE, lag.max = 100)
par(mfrow = c(1,1))

```

#

4. General correlation structure at low correlation. 

#

```{r, message=FALSE, warning=FALSE}
p=3
omega <- matrix(c(1,0,0,0,1,0,0,0,1),3,3)

A <- matrix(rnorm(p*p,mean=0,sd=1), p, p)
B <- A%*%t(A)
m <- max(eigen(B)$values)
phi0 <- B/(m+0.001)
Phi <- 0.50*phi0

scratch <- diag(p^2) - kronecker(Phi,Phi)
V.s <- solve(scratch)%*%vec(omega)
Tau0 <- matrix(V.s, nrow = p, byrow = TRUE)

# Markov Chain
chain <- as.matrix(mAr.sim(rep(0,p), Phi, omega, N = n))
aa <- ccf(chain[,1], chain[,2], type = "correlation", plot = TRUE, lag.max = 100)
bb <- ccf(chain[,1], chain[,3], type = "correlation", plot = TRUE, lag.max = 100)

```

#

5. General correlation structure at high correlation. 

#


```{r}
# Example 3: 3-dim, High correlation
p=3
omega <- matrix(c(1,0,0,0,1,0,0,0,1),3,3)

A <- matrix(rnorm(p*p,mean=0,sd=1), p, p)
B <- A%*%t(A)
m <- max(eigen(B)$values)
phi0 <- B/(m+0.001)
Phi <- 0.90*phi0

scratch <- diag(p^2) - kronecker(Phi,Phi)
V.s <- solve(scratch)%*%vec(omega)
Tau0 <- matrix(V.s, nrow = p, byrow = TRUE)

# Markov Chain
chain <- as.matrix(mAr.sim(rep(0,p), Phi, omega, N = n))
aa <- ccf(chain[,1], chain[,2], type = "correlation", plot = TRUE, lag.max = 100)
bb <- ccf(chain[,1], chain[,3], type = "correlation", plot = TRUE, lag.max = 100)

```
